\textbf{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\\Online Appendix

### 2018-07-02 redaction/dhp_travail.tex
Guillaume sur l’intro

\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
Universit\'{e} Paris-Dauphine, PSL Research University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author. Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE;  email: \url{lise.patureau@dauphine.fr} } }

%\date{July 2017\\PRELIMINARY AND INCOMPLETE \\ DO NOT QUOTE WITHOUT PERMISSION}

@@

\begin{abstract}
This paper investigates the role of additive costs in international transport costs, using information contained in the US imports flows database over 1974-2013. We find that the additive component of transport costs is sizeable, representing between 40 and 50\% of overall transport costs. Further, in the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects in the downward trend of international transport costs. In contrast to him, we find that trade composition effects do not matter much and when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs. Importantly, this difference of results can be attributed to the new way of modeling the per-unit (additive) component of transport costs we offer. In both aspects, our results point the importance of the additive component in accounting for international transport costs.
\end{abstract}

@@

Trade costs are central to international economic analysis. In particular, they are considered a major obstacle to international economic integration and international trade flows. Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000. Trade costs declined faster for more developed countries (\citealp{arvis2016trade}). Still, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few). Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries represent a 74\% markup over production costs.

Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}. In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs would represent a 21\% markup over production costs. Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows. If much of trade policy barriers have been removed over the second half of the twentieth century, this suggests that the transport costs component remains large and deserves attention. International transport costs accordingly stand at the heart of the paper.\smallskip

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}. Many argue that transport costs have substantially decreased with technological advance in transportation (\citealp{levinson2016box}), infrastructure development and new communication technologies (\citealp{Lafourcade_Thisse}). This view does not make consensus though. Distance as an incompressible cost is still a very important determinant of trade flows (\citealp{Disdier_Head08}); further, progress in transportation techniques (that reduces the constant-dollar cost of transport) should be compared with technical progress in production (that reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden. In accordance with this view, \cite{hummels2007} points out that transport costs in proportion of export prices have declined less quickly than expected.\smallskip

@@

Proving this point drives us to provide a careful modelling of transport costs. Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded. As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see, among others, \cite{Irrazabal_2015} or \citealp{martin2012}). Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs. This is obtained by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive costs. To do so, we exploit information contained in the US imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis. This database is the same as \cite{hummels2007}, extended over more recent years (2005-2013). \smallskip

Our results may be summarized in two main new findings. First, additive transport costs are quantitatively sizeable. On average over 1974-2013, the additive cost is estimated to be 1.8\% and 2.9\% of the export price in air and vessel transport respectively. This is slightly lower than our estimates for the ad-valorem component (3.5\% and 3.2\% for air/vessel respectively on average over the period). Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport mode. To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs. This represents a valuable insight for calibrating related international trade or business cycle models. We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs. Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor~2.

Second, we find that the main source of the downward trend in overall transport costs -by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013, does come from the reduction of transport cost ``per se''. Trade composition effects only have a limited influence, and when they matter (in maritime transport), they amplify the reduction of pure transport costs. Otherwise stated, pure transport costs have declined less than observed costs. This stands in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004). Our result thus deepens the paradox of the less-than-expected decline of transport costs. Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer. Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period. \smallskip

In both aspects, our results point the importance of the additive component in accounting for international transport costs. In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}). As underlined by \cite{alchian}, the relative price of two varieties of some good will depend on the level of additive trade costs. In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out''). This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity. Also, their estimates imply that doubling freight costs increases average free alongside (fas) export prices by 80 to 141 percent, consistent with high quality goods being sold in markets with high freight costs. Recent empirical studies based on micro-level data also provide a strong empirical support to the role of additive costs (i.e., cost per unit exported) in international trade costs. Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs. Using firm-level trade data \textbf{from which country?}, \cite{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs. \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework. This is confirmed by \cite{Irrazabal_2015}. While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains. One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citet{Irrazabal_2015}). One contribution of the paper is to fill this gap.}

@@

The paper is built as follows. First, we estimate the values of international transport costs annually throughout the period 1974-2013 (and by transport mode), explicitly distinguishing between the additive and the ad-valorem components. This is made in Section \ref{sec:data_method}, which reports the mean values of the transport costs estimated over the period (by transport mode). Relying on these estimation results, we then provide a decomposition of the transport costs trend patterns, between what comes from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs ``per se''. Section \ref{sec:results_trends} is devoted to this analysis. Section \ref{sec:conclu} concludes.

@@

Obviously, this dataset is not immune of limitations. First, it restricts our analysis to the study of international transport costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs. It is thus silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}. Second, in terms of transport costs \textit{per se}, our measure based on the cif-fas price gap captures the quantitative costs due to insurance, handling and freight, but it omits the other dimension of transport costs related to the time value of goods in transit. In this respect, this dataset embraces only a partial view of international transport barriers.
However, it is also clear that direct transport costs do represent a sizeable share of trade costs. According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...), that amount to a  9\% tax equivalent. Further, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers. They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).\footnote{In his survey, \citet{Hummels_1999} mentions several papers which all point that transport costs pose a barrier similar in size, or larger than tariffs. In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.}

@@

Our ambition to re-assess the underlying sources of the trend patterns of international transport costs requires a careful modeling of both the additive and the multiplicative components. The first step of the analysis then consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})). This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizeable component of the latter. These results constitute our first original contribution to the literature. We also provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation (see results in In Appendix \ref{app:diagnostic_test}).\smallskip

@@

Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs. On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to 10 cents (0.1 USD) per kilogram exported in maritime transport, and 1.12 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case. Expressed in terms of the export price, the additive costs amount to 2.9\% and 1.8 \% for ocean shipping and air transport respectively. This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\% and 2.5\% for vessel and air shipping respectively (on average over the period). Put differently, the additive component amounts to 48\% of total transport costs in Vessel, and 42\% in Air. This shows that the per-kg cost dimension is quantitatively sizeable. For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly a factor 2. Additionally, Appendix \ref{app:diagnostic_test} reports quality-of-fit diagnostic tests, that confirm the importance of the additive cost component. The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degree of freedom. An important aspect of these results is also to provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for more theoretical approaches needing calibration.\smallskip

Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question. Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of export price multiplied by the ad-valorem cost (expressed in Norwegian crowns), i.e. $\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and retaining their estimated weighted mean value, which is the one directly comparable to our own results). One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem). From this, we can rebuilt the ratio in similar terms to them, allowing us to gain in generality in this respect. Making use of our estimates for 2004 (see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail} in Appendix \ref{app:more_results}), we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.5\% in air transport. This may sound surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}. This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case. While the database of \citet{Irrazabal_2015} allows studying trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage of the two databases). In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment. \citet{Irrazabal_2015}'s results imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs. For the same export price multiplied by the ad-valorem cost of 100 USD, our results yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$. Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 47\% (in Vessel) and 25\% (in Air) and   additive trade costs are attributable to international transport. On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs.\smallskip

@@

In both transport modes, the additive component appears of sizeable importance throughout the period. Confirming the results obtained in Table \ref{tab:summary_results}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon. By contrast, they represent a sizeable and structural dimension of international transport costs. If the share of additive costs in total costs roughly amounts to 48\% (42\%) in Vessel (Air) transport on average, one can further notice that it substantially varies over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered. In the next section, we keep on investigating the importance of the (varying pattern of) additive costs, as regards with the determinants of transport costs time trends.

@@

This paper empirically studies the magnitude of additive costs in international transport costs, by exploiting the differences between the import and the export prices. Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2013, we estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean). Our results may be summarized in two main findings. First, we provide a quantitative measure of both the additive and the ad-valorem transport cost. We find that additive costs are quantitativeley sizeable: They amount to 2.8\% of the export price unit values for ocean shipping, 1.8\% for air transport. This is slighly lower that the estimated values for the d-valorem costs, around  2.5\% (vessel) and 3.2\% (air) as mean values over 1974-2013. Further, modeling the additive component significantly improves the quality of fit of the regression. Second, we show the importance of integrating the additive component in accounting for the time trend of international transport costs. Allowing for a varying share of additive costs in product/country/time dimension, we find that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role. If anything, trade composition effects have contributed to amplify the reduction in the pure transport costs, in particular in maritime transport. In both aspects, our results point the importance of the additive component in accounting for international transport costs.

>>>>

\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris-Dauphine, Université PSL, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author. Universit\'{e} Paris-Dauphine, Université PSL, LEDa, 75016 PARIS, FRANCE;  email: \url{lise.patureau@dauphine.fr} } }

@@

\begin{abstract}
This paper investigates the role of additive costs in international transport costs, using information contained in a US imports flows database over 1974-2013. We find that the additive component of transport costs is sizable, representing between 40 and 50\% of overall transport costs. Further, in the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects in the downward trend of international transport costs. In contrast to him, we find that trade composition effects do not matter much and, when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs. Importantly, this difference of results can be attributed to the new way of modeling the per-unit (additive) component of transport costs we offer. In both aspects, our results point to the importance of the additive component in accounting for international transport costs.
\end{abstract}

@@

Trade costs are central to international economic analysis. In particular, they are known to be a major obstacle to international economic integration and international trade flows. Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000. Trade costs declined faster for more developed countries (\citealp{arvis2016trade}). Still, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few). Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs represent a 74\% markup over production costs for industrialized countries.

Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariffs  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}. In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs represent a 21\% markup over production costs. Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows. If much of trade policy barriers have been removed over the second half of the twentieth century, this suggests that the transport costs component remains large and deserves attention. International transport costs accordingly stand at the heart of the paper.\smallskip

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}. Many argue that transport costs have substantially decreased with technological advance in transportation (\citealp{levinson2016box}), infrastructure development and new communication technologies (\citealp{Lafourcade_Thisse}). There is no consensus thought. Distance as an incompressible cost is still a very important determinant of trade flows (\citealp{Disdier_Head08}); further, progress in transportation techniques (that reduces the constant-dollar cost of transport) should be compared with technical progress in production (that reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden. In accordance with this view, \cite{hummels2007} points out that transport costs in proportion of export prices have declined more slowly than the common wisdom assumes.\smallskip

@@

Proving this point drives us to provide a careful modelling of transport costs. Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded. As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see, among others, \cite{Irrazabal_2015} or \citealp{martin2012}). Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs. We obtain this by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive costs. To do so, we exploit information contained in the Census Bureau's database of US imports flows over 1974-2013. This database contains a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis. This database is the same as \cite{hummels2007}, extended over more recent years (2005-2013). \smallskip

Our results may be summarized in two main new findings. First, additive transport costs are quantitatively sizable. On average over 1974-2013, the additive cost is estimated to be 1.8\% and 2.9\% of the export price in air and vessel transport respectively. This is slightly lower than our estimates for the ad-valorem component (3.5\% and 3.2\% for air/vessel respectively on average over the period). Put it differently, additive costs represent between 40 \% and 50\% of the overall transport costs, depending on the year and the transport mode. \textbf{GD Pourquoi est-ce qu'en faisant le calcul je trouve 33\% et 47,5\% ?} To the best of our knowledge, our paper is the first to provide such an extensive quantitative measure of both ad-valorem and additive costs in total transport costs. This represents a valuable insight for calibrating related international trade or business cycle models. We also provide an empirical assessment of what standard international trade models lose by skipping additive transport costs. Quantitatively, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor~2.

Second, we find that the main source of the downward trend in overall transport costs -by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013, comes mainly from the reduction of transport cost ``per se''. Trade composition effects only have a limited influence, and when they matter (in maritime transport), they amplify the reduction of pure transport costs. Otherwise stated, pure transport costs have declined less than observed costs. This stands in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004). Our result thus deepens the paradox of the less-than-expected decline of transport costs. Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer. Allowing the share of additive costs to vary across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period. \smallskip

In both aspects, our results point the importance of the additive component in accounting for international transport costs. In this respect, it contributes to the literature that challenges the dominant role of iceberg costs in international trade (even in the case of ice transport, see \citealp{bosker2018ice}). As shown by \cite{alchian}, the relative price of two varieties of some good will depend on the level of additive trade costs. In this context, the relative demand for more expensive/higher quality product goods increases with trade costs (``shipping the good apples out''). This is consistent with the findings by \citet{hummels_skiba}, who estimate the elasticity of freight rates with respect to price to be well below unity. Also, their estimates imply that doubling freight costs increases average free alongside (fas) export prices by 80 to 141 percent, consistent with high quality goods being sold in markets with high freight costs. Recent empirical studies based on micro-level data also provide a strong empirical support to the role of additive costs (i.e., cost per unit exported) in international trade costs. Based on a firm-product-level database of French exporters, \citet{martin2012} finds that firms charge higher unit values for exports to more remote countries, supporting the importance of additive costs. Using firm-level trade data from Colombia, \cite{Lashkaripour-2018} estimates that the transport cost elasticity varies greatly across industries, with a value of 0.8 for the average industry, in contrast with the iceberg specification.\footnote{Beyond the positive aspect, several recent papers also point out the normative implications of additive trade costs. \citet{sorensen2014} extends \citet{melitz}'s seminal model of international trade by including additive trade costs, in addition to the ad-valorem component. A key analytical result is that the welfare gain from a reduction in trade barriers is higher for a decrease in additive costs than a decrease in ad-valorem costs, due to the alteration of relative prices in a heterogeneous-firms trade framework. This is confirmed by \cite{Irrazabal_2015}. While these results suggest that important welfare gains can be achieved by reducing additive trade costs, not much progress has been done in quantifying such gains. One potential reason is the lack of an empirical characterization of the additive component of trade costs (the one exception being \citet{Irrazabal_2015}). One contribution of the paper is to fill this gap.}

@@

The paper is built as follows. First, we estimate the values of international transport costs annually throughout the period 1974-2013 by transport mode, explicitly distinguishing between the additive and the ad-valorem components.  Section \ref{sec:data_method} thus reports the mean values of the transport costs estimated over the period (by transport mode). Relying on these estimation results, we then provide a decomposition of the transport costs trend patterns, between what comes from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs ``per se''. Section \ref{sec:results_trends} is devoted to this analysis. Secion \ref{sec:robustness} checks its robustness to the separability assumption and the weighting scheme. Section \ref{sec:conclu} concludes.

@@

Obviously, this dataset is not immune of limitations. First, it restricts our analysis to the study of international transport costs, as our measure of the cif-fas price gap only covers freight, insurance and handling costs. It is thus silent about the others dimensions of international trade costs, unlike \citet{Irrazabal_2015}. Second, in terms of transport costs \textit{per se}, our measure based on the cif-fas price gap captures the quantitative costs due to insurance, handling and freight, but it omits the other dimension of transport costs related to the time value of goods in transit. In this respect, this dataset embraces only a partial view of international transport barriers.
However, it is also clear that direct transport costs do represent a sizable share of trade costs. According to \citet{anderson_wincoop_jel}, the 21\% markup over production costs coming from transport costs includes both directly measured freight costs (11\%) and ``indirect'' costs related to the time value of goods on their way to their export market (including holding cost for the goods in transit, inventory cost due to buffering the variability of delivery dates, preparation costs associated with shipment size...), that amount to a  9\% tax equivalent. Further, the evidence summarized in \citet{anderson_wincoop_jel} points to a persisting importance of direct transport costs, especially compared to other trade barriers. They remain more important than, e.g., policy barriers (8\% tax equivalent), language barrier (7\%) or information cost (6\%).\footnote{In his survey, \citet{Hummels_1999} mentions several papers which all point that transport costs pose a barrier similar in size, or larger than tariffs. In the same vein, \citet{limao_venables} highlight the importance of infrastructures for trade costs in general, through their impact on transport costs.}

@@

Our ambition to re-assess the underlying sources of the trend patterns of international transport costs requires a careful modeling of both the additive and the multiplicative components. The first step of the analysis then consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})). This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizable component of the latter. These results constitute our first original contribution to the literature. We also provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation (see results in In Appendix \ref{app:diagnostic_test}).\smallskip

@@

Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs. On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to 10 cents (0.1 USD) per kilogram exported in maritime transport, and 1.12 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case. Expressed in terms of the export price, the additive costs amount to 2.9\% and 1.8 \% for ocean shipping and air transport respectively. This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\% and 2.5\% for vessel and air shipping respectively (on average over the period). Put differently, the additive component amounts to 48\% of total transport costs in Vessel, and 42\% in Air. This shows that the per-kg cost dimension is quantitatively sizable. For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly a factor 2. Additionally, Appendix \ref{app:diagnostic_test} reports quality-of-fit diagnostic tests, that confirm the importance of the additive cost component. The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degree of freedom. An important aspect of these results is also to provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for more theoretical approaches needing calibration.\smallskip

Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question. Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of export price multiplied by the ad-valorem cost (expressed in Norwegian crowns), i.e. $\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and retaining their estimated weighted mean value, which is the one directly comparable to our own results). One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem). From this, we can rebuilt the ratio in similar terms to them, allowing us to gain in generality in this respect. Making use of our estimates for 2004 (see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail} in Appendix \ref{app:more_results}), we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.5\% in air transport. This may sound surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}. This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case. While the database of \citet{Irrazabal_2015} allows studying trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage of the two databases). In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment. \citet{Irrazabal_2015}'s results imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs. For the same export price multiplied by the ad-valorem cost of 100 USD, our results yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$. Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 47\% (in Vessel) and 25\% (in Air) and   additive trade costs are attributable to international transport. On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizable dimension of international transport costs.\smallskip

@@

In both transport modes, the additive component appears of sizable importance throughout the period. Confirming the results obtained in Table \ref{tab:summary_results}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon. By contrast, they represent a sizable and structural dimension of international transport costs. If the share of additive costs in total costs roughly amounts to 48\% (42\%) in Vessel (Air) transport on average, one can further notice that it substantially varies over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered. In the next section, we keep on investigating the importance of the (varying pattern of) additive costs, as regards with the determinants of transport costs time trends.

@@

This paper empirically studies the magnitude of additive costs in international transport costs, by exploiting the differences between the import and the export prices. Using SITC 3 and 4- digit cif-fas unit values taken from the US import database over 1974-2013, we estimate the two components (additive and ad-valorem) of transport costs, by transport mode (air or ocean). Our results may be summarized in two main findings. First, we provide a quantitative measure of both the additive and the ad-valorem transport cost. We find that additive costs are quantitativeley sizable: They amount to 2.8\% of the export price unit values for ocean shipping, 1.8\% for air transport. This is slighly lower that the estimated values for the d-valorem costs, around  2.5\% (vessel) and 3.2\% (air) as mean values over 1974-2013. Further, modeling the additive component significantly improves the quality of fit of the regression. Second, we show the importance of integrating the additive component in accounting for the time trend of international transport costs. Allowing for a varying share of additive costs in product/country/time dimension, we find that the decrease of international transport costs observed in the data is mostly attributable to a reduction in the \textit{pure} transport costs, with trade pattern composition effects playing a small role. If anything, trade composition effects have contributed to amplify the reduction in the pure transport costs, in particular in maritime transport. In both aspects, our results point the importance of the additive component in accounting for international transport costs.
